---
title: 'Data Analysis #2 Version 2 (75 points total)'
author: "huddleston, James"
output:
  word_document: default
  html_document: default
---

```{r setup, include = FALSE}
# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)

```

##Data Analysis #2

```{r analysis_setup1, message = FALSE, warning = FALSE}

# Perform the following steps to start the assignment.
 
# 1) Load/attach the following packages via library():  flux, ggplot2, gridExtra, moments, rockchalk, car.
# NOTE:  packages must be installed via install.packages() before they can be loaded.

library(flux)
library(ggplot2)
library(gridExtra)
library(moments)
library(rockchalk) # base R code replaces requirement for this package
library(car)

# 2) Use the "mydata.csv" file from Assignment #1 or use the file posted on the course site.  Reading
# the files into R will require sep = "" or sep = " " to format data properly.  Use str() to check file
# structure.

mydata <- read.csv("mydata.csv", sep = ",")
# mydata <- read.csv(file.path("c:...", "mydata.csv"), sep = ",")
# mydata <- read.csv(file.path("c:/Rabalone/", "mydata.csv"), sep = ",")

str(mydata)

```

### Test Items starts from here - There are 10 sections - total of 75 points ##############

##### Section 1: (5 points)

(1)(a) Form a histogram and QQ plot using RATIO. Calculate skewness and kurtosis using 'rockchalk.' Be aware that with 'rockchalk', the kurtosis value has 3.0 subtracted from it which differs from the 'moments' package. 

```{r Part_1a}
mydata$VOLUME <- mydata$LENGTH * mydata$DIAM * mydata$HEIGHT
mydata$RATIO  <- mydata$SHUCK / mydata$VOLUME
s  <- rockchalk::skewness(mydata$RATIO)
k  <- rockchalk::kurtosis(mydata$RATIO, excess = FALSE)
k2 <- rockchalk::kurtosis(mydata$RATIO)
par(mfrow = c(1, 2), oma = c(0, 0, 2, 0))
hist(mydata$RATIO, col = "steelblue", xlab = "Ratio",
     main = "Histogram of Ratio (Shuck / Vol)")
qqnorm(mydata$RATIO, main = "Q-Q Plot",
       col = ifelse(mydata$RATIO %in% c(boxplot.stats(mydata$RATIO)$out), 
                    "red", "steelblue"))
qqline(mydata$RATIO)
txt <- c("skewness:", round(s, 2), 
         "kurtosis:", round(k, 2), 
         "excess kurtosis:", round(k2, 2))
mtext(paste(txt, sep=" ", collapse="  "), outer = TRUE, cex = 1)
```

(1)(b) Tranform RATIO using *log10()* to create L_RATIO (Kabacoff Section 8.5.2, p. 199-200). Form a histogram and QQ plot using L_RATIO. Calculate the skewness and kurtosis. Create a boxplot of L_RATIO differentiated by CLASS.

```{r Part_1b}
mydata$L_RATIO <- log10(mydata$RATIO)
s  <- rockchalk::skewness(mydata$L_RATIO)
k  <- rockchalk::kurtosis(mydata$L_RATIO, excess = FALSE)
k2 <- rockchalk::kurtosis(mydata$L_RATIO)
par(mfrow = c(1, 3), oma = c(0, 0, 2, 0))
hist(mydata$L_RATIO, col = "steelblue", xlab = "Log Ratio",
     main = "Histogram of Log Ratio")
qqnorm(mydata$L_RATIO, main = "Q-Q Plot",
       col = ifelse(mydata$L_RATIO %in% c(boxplot.stats(mydata$L_RATIO)$out), 
                    "red", "steelblue"))
qqline(mydata$L_RATIO)
boxplot(mydata$L_RATIO ~ mydata$CLASS,
        main = "Boxplots of Log Ratio by Class",
        ylab = "Frequency",
        xlab = "Log Ratio Distribution by Class", 
        beside = TRUE,
        col = 'steelblue',
        pars = list(outcol = "red"))
txt <- c("skewness:", round(s, 2), 
         "kurtosis:", round(k, 2), 
         "excess kurtosis:", round(k2, 2))
mtext(paste(txt, sep=" ", collapse="  "), outer = TRUE, cex = 1)

```

(1)(c) Test the homogeneity of variance across classes using *bartlett.test()* (Kabacoff Section 9.2.2, p. 222). 

```{r Part_1c}
x <- bartlett.test(RATIO ~ CLASS, data = mydata)
y <- bartlett.test(L_RATIO ~ CLASS, data = mydata)
print(list('ratio:', x, 'log ratio:', y))

```

**Essay Question: Based on steps 1.a, 1.b and 1.c, which variable RATIO or L_RATIO exhibits better conformance to a normal distribution with homogeneous variances across age classes?  Why?** 

***Answer: (With homogeneous variances (Bartlett test) and normal distribution across age classes, L_RATIO displays better conformance then by failing to reject the null hypothesis. There is less skewness in the histogram and QQ plot and  evenly distributed outliers in the QQ plot and boxplots, and failing to reject the null hypothesis with the Bartlett test of homogeneity of variances, then L_RATIO shows better conformance to a normal distribution.)***


##### Section 2 (10 points) ###############################

(2)(a) Perform an analysis of variance with *aov()* on L_RATIO using CLASS and SEX as the independent variables (Kabacoff chapter 9, p. 212-229). Assume equal variances. Perform two analyses. First, fit a model with the interaction term CLASS:SEX. Then, fit a model without CLASS:SEX. Use *summary()* to obtain the analysis of variance tables (Kabacoff chapter 9, p. 227).

```{r Part_2a}

summary(aov(L_RATIO ~ CLASS + SEX + CLASS:SEX, data = mydata))
summary(aov(L_RATIO ~ CLASS + SEX, data = mydata))

```

**Essay Question:  Compare the two analyses.  What does the non-significant interaction term suggest about the relationship between L_RATIO and the factors CLASS and SEX?**

***Answer: (The Class and Sex are statistically significant, adding L_RATIO an intersection did not affect the significance. The interaction for Class and sex is insignificant by the p-value significance L_RATIO.)***

(2)(b) For the model without CLASS:SEX (i.e. an interaction term), obtain multiple comparisons with the *TukeyHSD()* function. Interpret the results at the 95% confidence level (*TukeyHSD()* will adjust for unequal sample sizes). 

```{r Part_2b}
TukeyHSD(aov(L_RATIO ~ CLASS + SEX, data = mydata))

```

**Additional Essay Question:  first, interpret the trend in coefficients across age classes. What is this indicating about L_RATIO?  Second, do these results suggest male and female abalones can be combined into a single category labeled as 'adults?' If not, why not?**

***Answer: (We Reject the null hypothesis that Infants and Male or Females are the same. If we are failing to reject the Males and Females then Males and Females can be combined into one Adult group.)***


######  Section 3: (10 points) ##################

(3)(a1) We combine "M" and "F" into a new level, "ADULT". (While this could be accomplished using *combineLevels()* from the 'rockchalk' package, we use base R code because many students do not have access to the rockchalk package.) This necessitated defining a new variable, TYPE, in mydata which had two levels:  "I" and "ADULT". 

```{r Part_3a1}
# here we show how to define the new variable TYPE using only base R functions (no need for outside packages)
mydata$TYPE <- character(nrow(mydata))  # initialize the TYPE column as all blanks
for (i in seq(along = mydata$SEX)) {
  mydata$TYPE[i] <- 'I'
  if (mydata$SEX[i] == 'M' || mydata$SEX[i] == 'F') mydata$TYPE[i] <- 'ADULT'
}
mydata$TYPE <- factor(mydata$TYPE)
cat('\nCheck on definition of TYPE object (should be an integer): ', typeof(mydata$TYPE))
cat('\nmydata$TYPE is treated as a factor: ', is.factor(mydata$TYPE), '\n')
table(mydata$SEX, mydata$TYPE)

```
(3)(a2)  Present side-by-side histograms of VOLUME. One should display infant volumes and, the other, adult volumes. 

```{r Part_3a2}
mydata$TYPE<-rockchalk::combineLevels(mydata$SEX, levs = c("F", "M"), "ADULT")
par(mfrow = c(1, 2))
hist(mydata$VOLUME[mydata$TYPE == 'I'], col = "steelblue", xlab = "Volume",
     main = "Infant Volumes", ylim = c(0, 150))
hist(mydata$VOLUME[mydata$TYPE == 'ADULT'], col = "indianred", xlab = "Volume",
     main = "Adult Volumes", ylim = c(0, 150))

```


**Essay Question: Compare the histograms.  How do the distributions differ? Are there going to be any difficulties separating infants from adults based on VOLUME?**

***Answer: (The Adult distribution is greater than 300 and the Infants are less than 300 is most obvious. The infants are skewed right and distributed with lower volumes. The adults are more evenly distributed-normally distributed. The infants have outliers more frequently. The volume is splitting the abalones.  )***

(3)(b) Create a scatterplot of SHUCK versus VOLUME and a scatterplot of their base ten logarithms, labeling the variables as L_SHUCK and L_VOLUME. Please be aware the variables, L_SHUCK and L_VOLUME, present the data as orders of magnitude (i.e. VOLUME = 100 = 10^2 becomes L_VOLUME = 2). Use color to differentiate CLASS in the plots. Repeat using color to differentiate by TYPE. 

```{r Part_3b}
L_SHUCK <- log10(mydata$SHUCK)
L_VOLUME <- log10(mydata$VOLUME)
mydata <- cbind(mydata,L_SHUCK,L_VOLUME)
plot(mydata$VOLUME, mydata$SHUCK, main="SHUCK by VOLUME", col=mydata$CLASS, xlab="VOLUME", ylab="SHUCK", cex=.8)
legend(0,160,unique(mydata$CLASS),col=1:length(mydata$CLASS), pch=1, bty="n", title="CLASS", cex=.8)
plot(mydata$L_VOLUME, mydata$L_SHUCK, main="Log 10 SHUCK by Log 10 VOLUME", col=mydata$CLASS, xlab="L_VOLUME", ylab="L_SHUCK", cex=.8)
legend(0.5,2.25,unique(mydata$CLASS),col=1:length(mydata$CLASS), pch=1, bty="n", title="CLASS", cex=.8)
plot(mydata$VOLUME, mydata$SHUCK, main="SHUCK by VOLUME", col=mydata$TYPE, xlab="VOLUME", ylab="SHUCK", cex=.8)
legend(0,160,unique(mydata$TYPE),col=1:length(mydata$TYPE), pch=1, bty="n", title="TYPE", cex=.8)
plot(mydata$L_VOLUME, mydata$L_SHUCK, main="Log 10 SHUCK by Log 10 VOLUME", col=mydata$TYPE, xlab="L_VOLUME", ylab="L_SHUCK", cex=.8)
legend(0.5,2.25,unique(mydata$TYPE),col=1:length(mydata$TYPE), pch=1, bty="n", title="TYPE", cex=.8)
```

**Additional Essay Question:  Compare the two scatterplots. What effect(s) does log-transformation appear to have on the variability present in the plot?  What are the implications for linear regression analysis? Where do the various CLASS levels appear in the plots? Where do the levels of TYPE appear in the plots?**

***Answer: (The log transformation gives more stability and linearity because without it the plot shows more scattered behavior with respect to classes and types where the volume is increasing. The log transformation allows for the distinction in class increasing as volume increases since the adults have a higher volume than the infants as seen in the cut between the A! and Infant abalones from the rest of the data.  The adults are to the right in the data. )***


######   Section 4: (5 points) ###################################

(4)(a1) Since abalone growth slows after class A3, infants in classes A4 and A5 are considered mature and candidates for harvest. Reclassify the infants in classes A4 and A5 as ADULTS. This reclassification could have been achieved using *combineLevels()*, but only on the abalones in classes A4 and A5. We will do this recoding of the TYPE variable using base R functions. We will use this recoded TYPE variable, in which the infants in A4 and A5 are reclassified as ADULTS, for the remainder of this data analysis assignment. 

```{r Part_4a1}
for (i in seq(along = mydata$TYPE)) {
  if (mydata$CLASS[i] == 'A4' || mydata$CLASS[i] == 'A5') mydata$TYPE[i] <- 'ADULT'
}
mydata$TYPE <- factor(mydata$TYPE)
cat('\nCheck on redefinition of TYPE object (should be an integer): ', typeof(mydata$TYPE))
cat('\nmydata$TYPE is treated as a factor: ', is.factor(mydata$TYPE), '\n')
cat('\nThree-way contingency table for SEX, CLASS, and TYPE:\n')
print(table(mydata$SEX, mydata$CLASS, mydata$TYPE))
```

(4)(a2) Regress L_SHUCK as the dependent variable on L_VOLUME, CLASS and TYPE (Kabacoff Section 8.2.4, p. 178-186, the Data Analysis Video #2 and Black Section 14.2). Use the multiple regression model: L_SHUCK ~ L_VOLUME + CLASS + TYPE. Apply *summary()* to the model object to produce results.

```{r Part_4a2}


index <- (mydata$CLASS == "A5")|(mydata$CLASS == "A4")
mydata$TYPE[index] <- combineLevels(mydata$TYPE[index], levs = c("I", "ADULT"), "ADULT")
model <- lm(L_SHUCK ~ L_VOLUME + CLASS + TYPE, data=mydata)
summary(model)

```

**Essay Question:  Interpret the trend in CLASS levelcoefficient estimates? (Hint:  this question is not asking if the estimates are statistically significant. It is asking for an interpretation of the pattern in these coefficients, and how this pattern relates to the earlier displays).**

***Answer: (As the Class Level increases the coefficients decreases, where the estimated coefficients decrease in L_SHUCK as the CLASS increases. there is a larger increase value in shuck when the abalones are younger in age. The Class level is larger and larger coefficient estimate  where the older the CLASS, the smaller the L_SHUCK.)***

**Additional Essay Question:  Is TYPE an important predictor in this regression? (Hint:  This question is not asking if TYPE is statistically significant, but rather how it compares to the other independent variables in terms of its contribution to predictions of L_SHUCK for harvesting decisions.)  Explain your conclusion.**

***Answer: (Type is not particularly great in this study for predicting the target values, but Volume and Clases A3, A4, and A5 are better in this regression. An example is the of L_SHUCK for harvesting decisions, type does not help much and because the value is not significant from the table values created. The type does not have a powerful coefficient nor is the p-value supporting that type is a predictor in this data for the case study here.  .)***

-----

The next two analysis steps involve an analysis of the residuals resulting from the regression model in (4)(a) (Kabacoff Section 8.2.4, p. 178-186, the Data Analysis Video #2).

-----

###### Section 5: (5 points) #################################

(5)(a) If "model" is the regression object, use model$residuals and construct a histogram and QQ plot. Compute the skewness and kurtosis. Be aware that with 'rockchalk,' the kurtosis value has 3.0 subtracted from it which differs from the 'moments' package. 

```{r Part_5a}
hist(model$residuals, main="Histogram of Residuals", col="dark grey", xlab="Residuals")
qqnorm(model$residuals, main="QQ Plot of Residuals", col="dark grey")
qqline(model$residuals)
skewness(model$residuals)
kurtosis(model$residuals)

```

(5)(b) Plot the residuals versus L_VOLUME, coloring the data points by CLASS and, a second time, coloring the data points by TYPE. Keep in mind the y-axis and x-axis may be disproportionate which will amplify the variability in the residuals. Present boxplots of the residuals differentiated by CLASS and TYPE (These four plots can be conveniently presented on one page using *par(mfrow..)* or *grid.arrange()*. Test the homogeneity of variance of the residuals across classes using *bartlett.test()* (Kabacoff Section 9.3.2, p. 222).  

```{r Part_5b}
par(mfrow = c(2,2))
plot(mydata$L_VOLUME, model$residuals, main="Residuals by Log 10 VOLUME", col=mydata$CLASS, xlab="L_VOLUME", ylab="Residuals", cex=.8, xlim=c(0.5,4))
legend(3.5,0.35,unique(mydata$CLASS),col=1:length(mydata$CLASS), pch=1, bty="n", title="CLASS", cex=.8)
plot(mydata$L_VOLUME, model$residuals, main="Residuals by Log 10 VOLUME", col=mydata$TYPE, xlab="L_VOLUME", ylab="Residuals", cex=.8, xlim=c(0.5,4))
legend(3.1,0.3,unique(mydata$TYPE),col=1:length(mydata$TYPE), pch=1, bty="n", title="TYPE", cex=.8)
boxplot(model$residuals ~ mydata$CLASS, main="Residuals by CLASS", ylab="Residuals")
boxplot(model$residuals ~ mydata$TYPE, main="Residuals by TYPE", ylab="Residuals")
par(mfrow = c(1,1))
bartlett.test(model$residuals ~ mydata$CLASS)
qchisq(.95,4)
```

**Essay Question:  What is revealed by the displays and calculations in (5)(a) and (5)(b)? Does the model 'fit'?  Does this analysis indicate that L_VOLUME, and ultimately VOLUME, might be useful for harvesting decisions? Discuss.**  

***Answer: (The histogram and QQnorm plot in 5a indicate the data is normally distributed. The residuals (in plots are clustered at volumes 2.5) fit the normal distribution, and the kurtosis is above zero indicating heavy tails with the skewness near zero, and small variance and small volumes for the Class A1 infants. The Bartlett's test, where the mean and variance residuals are even with the p-value showing that the null hypothesis can not be rejected, normally distributed. There is a large/heavy clustering in the right side of the graphs which is unusually non-characteristic based on the above explanation. (Last sentence I am trying to give a better answer))***

-----

There is a tradeoff faced in managing abalone harvest. The infant population must be protected since it represents future harvests. On the other hand, the harvest should be designed to be efficient with a yield to justify the effort. This assignment will use VOLUME to form binary decision rules to guide harvesting. If VOLUME is below a "cutoff" (i.e. a specified volume), that individual will not be harvested. If above, it will be harvested. Different rules are possible.

The next steps in the assignment will require consideration of the proportions of infants and adults harvested at different cutoffs. For this, similar "for-loops" will be used to compute the harvest proportions. These loops must use the same values for the constants min.v and delta and use the same statement "for(k in 1:10000)."  Otherwise, the resulting infant and adult proportions cannot be directly compared and plotted as requested. Note the example code supplied below.

-----

#### Section 6: (5 points) ########################

(6)(a) A series of volumes covering the range from minimum to maximum abalone volume will be used in a "for loop" to determine how the harvest proportions change as the "cutoff" changes. Code for doing this is provided.

```{r Part_6a}

idxi <- mydata$TYPE == "I"
idxa <- mydata$TYPE == "ADULT"

max.v <- max(mydata$VOLUME)
min.v <- min(mydata$VOLUME)
delta <- (max.v - min.v)/10000
prop.infants <- numeric(10000)
prop.adults <- numeric(10000)
volume.value <- numeric(10000)

total.infants <- sum(idxi)  
total.adults <- sum(idxa)

for (k in 1:10000) { 
	value <- min.v + k*delta
	volume.value[k] <- value
	prop.infants[k] <- sum(mydata$VOLUME[idxi] <= value)/total.infants
	prop.adults[k] <-  sum(mydata$VOLUME[idxa] <= value)/total.adults
}

# prop.infants shows the impact of increasing the volume cutoff for
# harvesting. The following code shows how to "split" the population at
# a 50% harvest of infants.

n.infants <- sum(prop.infants <= 0.5)
split.infants <- min.v + (n.infants + 0.5)*delta  # This estimates the desired volume.
split.infants

n.adults <- sum(prop.adults <= 0.5)
split.adults <- min.v + (n.adults + 0.5)*delta
split.adults
#

```

(6)(b) Present a plot showing the infant proportions and the adult proportions versus volume.value. Compute the 50% "split" volume.value for each and show on the plot.   

```{r Part_6b}

plot(volume.value,prop.adults, xlab="", ylab="", type="l", xlim=c(0,1000), ylim=c(0,1), col="red")
par(new=T)
plot(volume.value,prop.infants, xlab="", ylab="", type="l", axes=F, col="blue")
title(main="Proportion of Adults and Infants Protected", xlab="Volume", ylab="Proportion")
legend("bottomright", legend=c("Adult","Infant"), lty=1, col=c("red","blue"))
abline(h=.5)
abline(v=split.adults)
text(split.adults+33, .46, labels=round(split.adults,0), cex=.8)
abline(v=split.infants)
text(split.infants+33, .46, labels=round(split.infants,0), cex=.8)
```

**Essay Question:  The two 50% "split" values serve a descriptive purpose illustrating the difference between the populations. What do these values suggest regarding possible cutoffs for harvesting?** 

***Answer: (The values are a good cutoff in the splits. 50% of the Infants population are split and needing protection, then 134 is a limit. The cutoff for 50% for a volume is 384 for the lowest abalone's adults harvested. Meaning that the 50% is split between the infants and adults the histogram 3a displays this peak. The infants have small volumes compared to adults. The cutoff between the two splits results in both false positives and true negatives being lessened.)***

-----

This part will address the determination of a volume.value corresponding to the observed maximum difference in harvest percentages of adults and infants. To calculate this result, the vectors of proportions from item (6) must be used. These proportions must be converted from "not harvested" to "harvested" proportions by using (1 - prop.infants) for infants, and (1 - prop.adults) for adults. The reason the proportion for infants drops sooner than adults is that infants are maturing and becoming adults with larger volumes.

-----

###### Section 7: (10 points)  #######################

(7)(a) Evaluate a plot of the difference ((1 - prop.adults) - (1 - prop.infants)) versus volume.value. Compare to the 50% "split" points determined in (6)(a). There is considerable variability present in the peak area of this plot. The observed "peak" difference may not be the best representation of the data. One solution is to smooth the data to determine a more representative estimate of the maximum difference.

```{r Part_7a}
propharvestinfants <- 1-prop.infants
propharvestadults <- 1-prop.adults
propharvestdiff <- propharvestadults-propharvestinfants
plot(volume.value, propharvestdiff, type="l", main="Difference in Harvest Proportions", col="purple", xlab="Volume", ylab="Difference")


```

(7)(b) Since curve smoothing is not studied in this course, code is supplied below. Execute the following code to create a smoothed curve to append to the plot in (a). The procedure is to individually smooth (1-prop.adults) and (1-prop.infants) before determining an estimate of the maximum difference. 

```{r Part_7b}

y.loess.a <- loess(1 - prop.adults ~ volume.value, span = 0.25,
	family = c("symmetric"))
y.loess.i <- loess(1 - prop.infants ~ volume.value, span = 0.25,
	family = c("symmetric"))
smooth.difference <- predict(y.loess.a) - predict(y.loess.i)

```

(7)(c) Present a plot of the difference ((1 - prop.adults) - (1 - prop.infants)) versus volume.value with the variable smooth.difference superimposed. Determine the volume.value corresponding to the maximum smoothed difference (Hint:  use *which.max()*). Show the estimated peak location corresponding to the cutoff determined.

```{r Part_7c}
plot(volume.value, propharvestdiff, type="l", main="Difference in Harvest Proportions", col="purple", xlab="Volume", ylab="Difference")
lines(smooth.difference, type="l", col="dark grey")
abline(v=volume.value[which.max(smooth.difference)], col="dark grey", lty=2)
cutoff1 <- volume.value[which.max(smooth.difference)]
text(volume.value[which.max(smooth.difference)], propharvestdiff[which.max(smooth.difference)], label=round(volume.value[which.max(smooth.difference)],2), cex=.8, pos=1)


```

(7)(d) What separate harvest proportions for infants and adults would result if this cutoff is used? Show the separate harvest proportions (NOTE:  the adult harvest proportion is the "true positive rate" and the infant harvest proportion is the "false positive rate").

Code for calculating the adult harvest proportion is provided.

```{r Part_7d}
(1 - prop.adults)[which.max(smooth.difference)]  # [1] 0.7416332
propharvestadults[which.max(smooth.difference)]
propharvestinfants[which.max(smooth.difference)]
```

-----

There are alternative ways to determine cutoffs. Two such cutoffs are described below.

-----

######  Section 8: (10 points)  ###################

(8)(a) Harvesting of infants in CLASS "A1" must be minimized. The smallest volume.value cutoff that produces a zero harvest of infants from CLASS "A1" may be used as a baseline for comparison with larger cutoffs. Any smaller cutoff would result in harvesting infants from CLASS "A1."  

Compute this cutoff, and the proportions of infants and adults with VOLUME exceeding this cutoff. Code for determining this cutoff is provided. Show these proportions.

```{r Part_8a}
volume.value[volume.value > max(mydata[mydata$CLASS == "A1" & mydata$TYPE == "I", "VOLUME"])][1]
cutoff2 <- volume.value[volume.value > max(mydata[mydata$CLASS == "A1" & mydata$TYPE == "I", "VOLUME"])][1]
propharvestinfants[which(volume.value==cutoff2)]
propharvestadults[which(volume.value==cutoff2)]
```

(8)(b) Another cutoff is one for which the proportion of adults not harvested equals the proportion of infants harvested. This cutoff would equate these rates; effectively, our two errors:  'missed' adults and wrongly-harvested infants. This leaves for discussion which is the greater loss:  a larger proportion of adults not harvested or infants harvested?  This cutoff is 237.7383. Calculate the separate harvest proportions for infants and adults using this cutoff. Show these proportions.  Code for determining this cutoff is provided.  

```{r Part_8b}
volume.value[which.min(abs(prop.adults - (1-prop.infants)))] # [1] 237.6391
cutoff3 <- volume.value[which.min(abs(prop.adults - (1-prop.infants)))]
propharvestinfants[which(volume.value==cutoff3)]
propharvestadults[which(volume.value==cutoff3)]
```


##### Section 9: (5 points) ###########

(9)(a) Construct an ROC curve by plotting (1 - prop.adults) versus (1 - prop.infants). Each point which appears corresponds to a particular volume.value. Show the location of the cutoffs determined in (7) and (8) on this plot and label each. 

```{r Part_9}
plot(propharvestinfants, propharvestadults, xlab="Infant Harvest Proportion", ylab="Adult Harvest Proportion", main="ROC Curve of Adult and Infant Harvest Proportions", type="l", col="dark green")
points(propharvestinfants[which(volume.value==cutoff1)], propharvestadults[which(volume.value==cutoff1)])
text(.3, .7, labels="max difference vol=261", cex=.8)
points(propharvestinfants[which(volume.value==cutoff2)], propharvestadults[which(volume.value==cutoff2)])
text(.42, .82, labels="zero A1 inf vol=206", cex=.8)
points(propharvestinfants[which(volume.value==cutoff3)], propharvestadults[which(volume.value==cutoff3)])
text(.35, .75, labels="equal harvest vol=237", cex=.8)

```

(9)(b) Numerically integrate the area under the ROC curve and report your result. This is most easily done with the *auc()* function from the "flux" package.   Areas-under-curve, or AUCs, greater than 0.8 are taken to indicate good discrimination potential. 

```{r Part_9b}
auc(propharvestinfants, propharvestadults)

```


##### Section 10: (10 points) ###################

(10)(a) Prepare a table showing each cutoff along with the following:
 	1) true positive rate (1-prop.adults,
 	2) false positive rate (1-prop.infants),
 	3) harvest proportion of the total population
 	
```{r Part_10} 	
table <- matrix(c(cutoff1, propharvestadults[which.max(smooth.difference)], propharvestinfants[which.max(smooth.difference)], sum(mydata$VOLUME>cutoff1)/1036, cutoff2, propharvestadults[which(volume.value==cutoff2)], propharvestinfants[which(volume.value==cutoff2)], sum(mydata$VOLUME>cutoff2)/1036, cutoff3, propharvestadults[which(volume.value==cutoff3)], propharvestinfants[which(volume.value==cutoff3)], sum(mydata$VOLUME>cutoff3)/1036), ncol=4, byrow=TRUE)
colnames(table) <- c("Volume", "TPR", "FPR", "Prop Yield")
rownames(table) <- c("max difference", "zero A1", "equal harvest")
table <- as.table(table)
table

```
 	
**Essay Question: Based on the ROC curve, it is evident a wide range of possible "cutoffs" exist. Compare and discuss the three cutoffs determined in this assignment.**   

***Answer: (The lowest proportional yield is the max difference cutoff. The highest proportional yield and positive rate is the zero harvest. The equal is the zero-harvest group or equal error between these two-equal harvest-the second-best harvest. The biggest volume and smallest proportion of total abalones cutoff is the difference in the adult and infant proportions. If Class 1A infants have a lower cutoff and the larger true positive rate and proportion then the higher false positive rate.  )***



**Final Essay Question:  Assume you are expected to make a presentation of your analysis to the investigators How would you do so?  Consider the following in your answer:**

1. Would you make a specific recommendation or outline various choices and tradeoffs?
2. What qualifications or limitations would you present regarding your analysis?
3. If it is necessary to proceed based on the current analysis, what suggestions would you have for implementation of a cutoff?  4)  What suggestions would you have for planning future abalone studies of this type? 

***Answer: (The recommendation is positive and negative approach to options in the table. I would try not to be biased in persuading the researchers in a particular way. Showing an example for improvement and as positive and showing more than one table of data comparison with more data samples from the study as improvement.
The researches need to see the plots created here, with the outliers and challenges that exist in interpreting the data with limited knowledge about gathering and using the data to make decisions. The plots and outliers: other techniques to use in analysis are regression equation and normality of the residuals because it is difficult to determine sex and bias may play a part in analysis and decision making that effect the overall results based on data collection and recording data. 
The cutoff method for a sample area to prevent overharvesting. Testing more and adding a control group. Determining an alternative approach for comparison on data gathering and reviewing data for analysis.
Are the regions comparable where the data is taken, do all data gathered represent the population of all the abalones? The ecosystem that the animals reside in different regions on temperature, seasons being longer, more food available in some area and man (human) effects-pollution.)***